Uncertainty Estimation Methods for Countering Attacks on Machine-Generated Text Detectors

### 2025-05-09 paper/main.tex
add RAID results

@@

@@

We got the embeddings for all texts from both labels using the LLM (Llama-3.1-8B-Instruct) and utilized contextual logits to compute four metrics: perplexity, mean token entropy, Monte Carlo sequence entropy and Mahalanobis distance. The results can be seen in Table 1. Some metrics show pronounced variation between human- and machine-generated texts. This finding is promising, as it suggests that the distinction between MG and HW text distributions can potentially be detected using uncertainty estimation (UE). In the following sections, we explore this idea further to develop a more robust detection framework.

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{ROC-AUC} & \textbf{Train Time (s)} \\
\hline
BERT Classifier & 0.9942 & 0.9954 & 1489.0528 \\
Neural Classifier with uncertainty & 0.8183 & 0.7942 & 208.9576 \\
Random forest with uncertainty & 0.8103 & 0.7831 & 6.7727 \\
Logistic Regression with uncertainty & 0.7744 & 0.7317 & 0.0134 \\
\hline
\end{tabular}
\caption{Performance comparison of classification approaches with uncertainty estimation on data from arXiv}
\label{tab:model-performance}
\end{table}

Our results demonstrate that uncertainty estimation methods can effectively identify machine-generated texts, with our proposed model achieving good ROC-AUC performance (0.783) while requiring minimal training time (<10 seconds). The research potential in this field is substantial—with greater computational resources, we could expand uncertainty estimation through ensemble methods and analyze more diverse text sources, paving the way for next-generation detectors that combine high accuracy, low computational costs, and resilience against evolving generation techniques.

>>>>

\subsection{M4GT}

@@

We got the embeddings for all texts from both labels using the LLM (Llama-3.1-8B-Instruct) and utilized contextual logits to compute four metrics: perplexity, mean token entropy, Monte Carlo sequence entropy and Mahalanobis distance.

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{ROC-AUC} & \textbf{Train Time (s)} \\
\hline
BERT Classifier & 0.9942 & 0.9954 & 1489.0528 \\
Neural Classifier with uncertainty & 0.8183 & 0.7942 & 208.9576 \\
Random forest with uncertainty & 0.8103 & 0.7831 & 6.7727 \\
Logistic Regression with uncertainty & 0.7744 & 0.7317 & 0.0134 \\
\hline
\end{tabular}
\caption{Performance comparison of classification approaches with uncertainty estimation on arXiv data from M4GT}
\label{tab:model-performance}
\end{table}

\subsection{RAID}

We now proceed to the primary objective of our study - investigating the performance of detection metrics on the RAID attack dataset. For this purpose, we selected text samples from the Reddit domain, ensuring uniform distribution across different generative models. The current model list includes: ChatHPT, GPT-2, GPT-3, GPT-4, MPT, MPT-Chat, LLaMA-Chat, Cohere, Cohere-Chat, Mistral, Mistral-Chat, and human-authored texts.

Unlike the M4GT benchmark, our dataset introduces a new critical feature - attack type. The texts are uniformly distributed across both these features (model origin and attack type), ensuring approximately equal representation of texts generated by model M with attack type A. The final dataset comprises approximately 30,000 text samples, with 20\% allocated to the test set and the remaining 80\% to the training set.

We maintained the original classification models without modifications, and obtained the following experimental results:

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{ROC-AUC} & \textbf{Train Time (s)} \\
\hline
BERT Classifier & 0.9538 & 0.9532 & 2362.1725 \\
Neural Classifier with uncertainty & 0.8987 & 0.8977 & 378.1838 \\
Random Forest with uncertainty & 0.8992 & 0.8987 & 10.7419 \\
Logistic Regression with uncertainty & 0.7271 & 0.7258 & 0.0350 \\
\hline
\end{tabular}
\caption{Performance comparison of classification approaches with uncertainty estimation on Reddit data from RAID}
\label{tab:model-performance-comparison}
\end{table}

Our experimental results reveal three key findings: (1) the BERT classifier experienced significant accuracy degradation (dropping from 0.9942 to 0.9538) compared to its performance on the M4GT dataset, likely due to adversarial text manipulations; (2) our uncertainty-enhanced models demonstrate substantially improved performance, becoming competitive with the LLM classifier; and (3) while maintaining significantly faster training times, these uncertainty-based approaches now achieve comparable accuracy and ROC-AUC scores, trailing the LLM classifier by just 0.06, suggesting they offer an effective balance between computational efficiency and detection performance in adversarial settings.

Our investigation extended to evaluating the models' capability for multiclass classification, which presents a more challenging task. The experimental results revealed the following key findings:

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{ROC-AUC (ovr)} & \textbf{Train Time (s)} \\
\hline
BERT Classifier & 0.6689 & 0.9104 & 2514.6458 \\
Neural Classifier with uncertainty & 0.6306 & 0.8786 & 418.5270 \\
Random Forest with uncertainty & 0.5799 & 0.8646 & 25.0479 \\
Logistic Regression with uncertainty & 0.5654 & 0.8046 & 1.2399 \\
\hline
\end{tabular}
\caption{Multiclass classification performance comparison on adversarial text data}
\label{tab:model-performance-comparison}
\end{table}

The experimental results demonstrate that all models exhibit competent performance in the multiclass classification task, indicating their strong capability to effectively discriminate between different generative sources.

Our results demonstrate that uncertainty estimation methods can effectively identify machine-generated texts, with our proposed model achieving good ROC-AUC performance (0.899) while requiring minimal training time (10 seconds for Random Forest and 378 seconds for Neural Classifier). The research potential in this field is substantial—with greater computational resources, we could expand uncertainty estimation through ensemble methods and analyze more diverse text sources, paving the way for next-generation detectors that combine high accuracy, low computational costs, and resilience against evolving generation techniques.

\section{Appendix}

The statistical metrics presented in Table 4 (pertaining to the first subset of the M4GT dataset) reveal significant variations in U) across different class labels. This observed variability in UE values provides a discriminative feature space that enables effective separation between text samples originating from different generative sources.

@@
